## Exercises

### 1. Conditionals

Consider the vector:

```{r}
grades <- c(85, 77, 90, 65, 89, 72, 91, 68, 80)
```

```{r}
# Creates an empty vector called grade_tags
grade_tags <- c()

# Initializes the vector with the conditional values
for (i in seq_along(grades)) {
    if (grades[i] > 90) {
        grade_tags[i] <- "Excellent"
    } else if (grades[i] >= 80) {
        grade_tags[i] <- "Good"
    } else {
        grade_tags[i] <- "Fair"
    }
}

print(grade_tags)
```


Write a conditional script that:

1. Tags grades above 90 as "Excellent"
2. Between 80 and 90 (inclusive) as "Good"
3. Below 80 as "Fair"

Store the result in a new vector called grade_tags. Ensure the order in grade_tags corresponds to the order of grades in the grades vector.


### 2. For

Create a sequence of numbers from 3 to 15. Using a for loop, compute the cumulative product of these numbers.

```{r}
sequence <- c()
totalproduct <- 1
i <- 1
for(val in 3:15) {
    sequence[i] <- val
    totalproduct <- totalproduct * sequence[i]
    i <- i + 1
}
print(totalproduct)
```
### 3. While

Starting with a number 50, decrease it by 5% in every iteration using a while loop. Continue the iterations until the number goes below 20. Return the number of iterations it took.

```{r}
iteration <- 0
count <- 50
while(count > 20) {
    count <- count * 0.95
    iteration <- iteration + 1
}
print(iteration)
```

### 4. Functions with conditionals and loops

Write a function named series_sum that accepts two arguments: a start value 's' and an end value 'e'. The function should sum all numbers from 's' to 'e'. However, if a number in the series is divisible by 3, it should be skipped.

**Tip**: The remainder operator `%%` returns the remainder of the division of two numbers:
 
 ```{r}
total_non_three_sum <- 0
series_sum <- function(s, e) {
    for(val in s:e) {
        if (val %% 3 != 0) {
            total_non_three_sum <- total_non_three_sum + val
        }
    }
    print(total_non_three_sum)
}

# Unit Tests
series_sum(2, 5)
series_sum(5, 2)

 ```
 
### 5. Recursive Functions

The Lucas series is a sequence of numbers similar to the Fibonacci series but starts with 2 and 1 instead of 0 and 1. Write a recursive function lucas that computes the nth value of the Lucas series.
 
```{r}
 lucas <- function(n) {
    if (n == 0) {
        return(2)
    } else if (n == 1) {
        return(1)
    } else {
       return(lucas(n-2) + lucas(n-1))
    }
 }

print(lucas(10))

```
