1. **Data Parsing and Aggregation:**
    * Read the contaminant and ecological community data from the CSV files.
    * Parse the data, ensuring that any inconsistencies, missing values, or anomalies are handled.
    * Aggregate data so that it can be efficiently analyzed. This might involve merging data by date and location, calculating averages, or creating summarized tables.

2. **Calculation of Diversity Indexes:**
    * For each ecological community data set, calculate a diversity index (e.g., Shannon index, Simpson’s diversity index) over time.
    * Visualize the diversity indexes on a timeline for each location.

3. **Relate Contaminant Data to Diversity Indexes:**
    * Analyze the relationship between contaminant levels and the diversity indexes. 
    * Look for correlations, trends, or patterns that might suggest that certain contaminants have significant impacts on ecological diversity.

1. Start by creating a function to read and organize the 
contaminants data. your function should receive as an argument 
the a path to a directory containing the contamination data. 
It should then read these files (the `dir()` function can help 
to find the correct files) and merge them into a single data.frame. 
The function should return a data.frame with the following columns: 
`Date`, `Mercury`, `Lead`, and `Location`.

```{r}
library(tidyverse)

# Function that takes in a directory path and returns a data frame 
# containing all information in the csv files in that path
# more specifically using date, mercury, lead, and location as the 
# headers

all_contaminants_df <- function(directory_path) {
    # Stores the files in a list
    files <- dir(directory_path, pattern = ".*csv", full.names = TRUE)

    # Creates an empty Data Frame
    contaminants_df <- data.frame(
        Date = character(),
        Mercury = numeric(),
        Lead = numeric(),
        Location = numeric()
    )

    # Reads the files
    for (file in files) {
        df <- read.csv(file) 
        # Appends the information in each file into the df
        contaminants_df <- rbind(contaminants_df, df)
    }
    return(contaminants_df)
}

# Relative Path since the files are contained within the same folder
directory_path <- "contaminants"
contaminants <- all_contaminants_df(directory_path)
contaminants

```

Alternative:
```{r}
library(tidyverse)

path <- "contaminants"
readData <- function(path) {
    read_csv(dir(path, full.names = TRUE))
}
readData(path)

```


2. With all the contaminants data in a single data.frame, 
**summarize** this data set by calculating the average concentration 
per month (That is, the average in 2010-01, 2010-02, ...) 
for each of the two contaminants.

```{r}
library(stringr)

# Generates a new column called months using REGEX to extract dates
contaminants$Month <- format(as.Date(contaminants$Date, format = "%Y-%m-%d"), "%Y-%m")

# Generates a new data frame containing the average mercy and 
# lead concentrations 
avg_monthly_concentration <- contaminants %>% 
    group_by(Month, Location) %>% 
    summarise(
        avg_mercury_concentration = mean(Mercury, na.rm = TRUE),
        avg_lead_concentration = mean(Lead, na.rm = TRUE)
    )

avg_monthly_concentration

```

3. Load the species count data into R, wrangle it however 
you see fit, there are several possibilities. You may use the data 
in the format it is in or modify it using the tidy data principles. 

```{r}
library(readr)

# Regular Table with an additional column called total 
# containing the observation total 
species <- read.table("species_counts.csv", header = TRUE, sep = ",")
species

```

# 4. Next, modify the `Simpson()` and `Shannon()` diversity index 
# functions we built in class to handle the data in the format you 
# created in the previous step. Calculate the diversity index for 
# each observation of the two communities.

```{r}
getProb = function(x, na.rm = FALSE) {
    x/sum(x, na.rm = TRUE)
}

Simpson <- function(x, na.rm = FALSE){
  p = getProb(x)
  D = sum(p^2, na.rm = TRUE)
  return(c(Simpson = D))
}

Shannon <- function(x, na.rm = FALSE){
  p = getProb(x)
  H = -sum(p*log(p), na.rm = TRUE)
  return(c(Shannon = H))
}

diversity <- subset(species[,1:2]) %>% 
    mutate(Simpson = apply(species[,3:12], 1, Simpson)) %>% 
    mutate(Shannon = apply(species[,3:12], 1, Shannon))
diversity

```

5. Finally, join the two datasets (diversity measures and 
average contaminants) into a single data.frame.  
(Hint: use one of the join functions)
```{r}
combined <- left_join(diversity, avg_monthly_concentration, by = c("Month", "Location")) %>%
    arrange(Month, Location)
combined

```

6. Using this combined data set make plots, measure correlations, 
and look for patterns. For example: 
How do the contaminants behave in both locations? 
How does diversity change across time and locations? 
Are the two related? 
(Hint: see the first precept for a simple scatter plot code, 
and see the `cor()` function)

```{r}
library(dplyr)
library(ggplot2)

Month <- combined$Month
Simpson_Indexes <- combined$Simpson
Shannon_Indexes <- combined$Shannon
Location <- combined$Location
Lead <- combined$avg_lead_concentration
Mercury <- combined$avg_mercury_concentration


# 2 plots: ggplot & barplot representing Simpson Index vs Time (Months)
# ggplot separated by location
simpson_ggplot <- ggplot(
    data = combined, 
        aes(
            x = Month, 
            y = Simpson,
            fill = Location
        )
    ) + geom_col(width = 0.5) +
    theme(axis.text.x = element_text(angle = 90))

# barplot representing Simpson Index vs Time
barplot(combined$Simpson,
    names.arg = combined$Month,
    xlab = "Month", ylab = "Simpson",
    main = "Bar Plot of Month vs. Simpson",
    col=rgb(0.2,0.4,0.6,0.6),
)   

# A scatterplot representing the Simpson indexes, but separated
# by location, (wasn't able to draw a trend line) 
simpson_scatter <- ggplot(
    data = combined, 
        aes(
            x = Month, 
            y = Simpson,
            fill = Location
        )
    ) + theme(axis.text.x = element_text(angle = 90)) +
    geom_point(aes(colour=Location)) + 
    geom_smooth(method = "lm", formula = y ~ x) +
    facet_wrap(~ Location, ncol = 1)
simpson_scatter

# 2 plots: ggplot & barplot representing Shannon Index vs Time (Months)
# ggplot separates by location
shannon_ggplot <- ggplot(
    data = combined, 
        aes(
            x = Month, 
            y = Shannon,
            fill = Location
        )
    ) + geom_col(width = 0.5) +
    theme(axis.text.x = element_text(angle = 90))

# barplot representing Simpson Index vs Time
barplot(combined$Shannon,
    names.arg = combined$Month,
    xlab = "Month", ylab = "Shannon",
    main = "Bar Plot of Month vs. Shannon",
    col=rgb(0.2,0.4,0.6,0.6),
)  

# A scatterplot representing the shannon indexes, but separated
# by location, (wasn't able to draw a trend line)
shannon_scatter <- ggplot(
    data = combined, 
        aes(
            x = Month, 
            y = Shannon,
            fill = Location
        )
    ) + theme(axis.text.x = element_text(angle = 90)) +
    geom_point(aes(colour=Location)) + 
    facet_wrap(~ Location, ncol = 1)
shannon_scatter


# Analyzing contaminants by Location
# A scatterplot representing the lead concentrations 
# in each location over time
lead <- ggplot(
    data = combined, 
        aes(
            x = Month,
            y = Lead,
            fill = Location
        )
    ) + theme(axis.text.x = element_text(angle = 90)) +
    geom_point(aes(colour=Location)) + 
    facet_wrap(~ Location, ncol = 1)

# A scatterplot representing the lead concentrations 
# in each location over time
mercury <- ggplot(
    data = combined, 
        aes(
            x = Month,
            y = Mercury,
            fill = Location
        )
    ) + theme(axis.text.x = element_text(angle = 90)) +
    geom_point(aes(colour=Location)) + 
    facet_wrap(~ Location, ncol = 1)

```
# Measure Correlations
```{r}

correlations <- data.frame(Location = character(),
                           Simpson = numeric(),
                           Shannon = numeric(),
                           Lead = numeric(),
                           Mercury = numeric())

# For River location
river_data <- subset(combined, Location == "River")
simpson_cor_river <- cor(river_data$Simpson, c(1:30))
shannon_cor_river <- cor(river_data$Shannon, c(1:30))
lead_cor_river <- cor(river_data$avg_lead_concentration, c(1:30))
mercury_cor_river <- cor(river_data$avg_mercury_concentration, c(1:30))

correlations <- rbind(correlations, 
    data.frame(
        Location = "River",
        Simpson = simpson_cor_river,
        Shannon = shannon_cor_river,
        Lead = lead_cor_river,
        Mercury = mercury_cor_river
        )
    )

# For Lake location
lake_data <- subset(combined, Location == "Lake")
simpson_cor_lake <- cor(lake_data$Simpson, c(1:30))
shannon_cor_lake <- cor(lake_data$Shannon, c(1:30))
lead_cor_lake <- cor(lake_data$avg_lead_concentration, c(1:30))
mercury_cor_lake <- cor(lake_data$avg_mercury_concentration, c(1:30))

correlations <- rbind(correlations, 
    data.frame(
        Location = "Lake",
        Simpson = simpson_cor_lake,
        Shannon = shannon_cor_lake,
        Lead = lead_cor_lake,
        Mercury = mercury_cor_lake
        )
    )

correlations

```
Observations of patterns

7. Write a single paragraph that summarizes your conclusions.

Conclusions:
Based on the graphs and correlations, it seems that both
mercury and lead concentrations in each location have a
positive correlation over time, meaning that their average 
monthly concentrations are increasing in each location. 
With respect to this, the simpson diversity index has 
a positive correlation with respect to time with the lake location
having a much higher correlation than that of the river location.
From the graphs, we can see that the Simpson indexes increase over
time meaning that the probability that two individuals sampled from 
the same population will be from the same species is higher.
The Shannon indexes hold a negative correlation and the scatterplots
show that the indexes decrease over time. This indicates that 
the diversity in each location in decreasing. The likely
conclusion from this data is that as the mercury and lead
concentrations increase in each community over time is correlated
with the diversity of the species in each community decrease 
over time
